---
title: "The Decade of Disruption: US-China Tech Giants Comparison"
subtitle: "A Data Story of Rise, Resilience and Realignment (2015-2025)"
date: "2025-11-20"
format:
  html:
    # 添加以下行来消除所有消息和警告
    theme: flatly
    toc: true
    toc-depth: 3
    toc-location: left
    code-fold: true
    code-summary: "Show Code"
    css: styles.css
    echo: false
    warning: false
    message: false
editor: 
  markdown: 
    wrap: 72
---

```{r}
# 在setup块开头添加，消除冲突警告
library(conflicted)
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")
library(tidyverse)
library(plotly)
library(lubridate)
library(scales)
library(here)


cleaned_data <- read_csv(here("data", "processed", "stock_prices_cleaned.csv"))
source(here("scripts","03_visualization.R"))
```

```{r}
knitr::opts_chunk$set(

echo = FALSE,

warning = FALSE,

message = FALSE,

fig.align = "center",

fig.width = 10,

fig.height = 6

)
```

## Executive Summary

Data analysis reveals distinct developmental trajectories between US and
Chinese tech giants over the past decade. US technology companies
(represented by NVDA and MSFT) have demonstrated sustained growth driven
by technological innovation and global market expansion, while Chinese
companies have shown stronger risk resilience and adaptability to
domestic market conditions. Notably, the AI technological revolution and
geopolitical factors have emerged as critical variables shaping market
dynamics over the past years.

```{r}
stats_summary <- cleaned_data %>%
  group_by(symbol, region) %>%
  summarise(
    total_return = (last(adjusted) - first(adjusted)) / first(adjusted),
    .groups = 'drop'
  )

ggplot(stats_summary, aes(x = reorder(symbol, total_return), y = total_return, fill = region)) +
  geom_col() +
  scale_y_continuous(labels = percent_format()) +
  coord_flip() +
  labs(
    title = "Total Returns: 2015-2025",
    subtitle = "Cumulative performance across all companies",
    x = "", y = "Total Return", fill = "Region"
  ) +
  scale_fill_manual(values = c("US" = "#3498db", "China" = "#e74c3c")) +
  theme_minimal()
```

## Chapter 1: The Starting Line (2015)

**2015 Market Structure**:

**US Dominance**: Apple, Microsoft and others had established solid
ecosystem advantages

**China's Rising Presence**: Alibaba and Tencent experienced rapid
growth during the mobile internet wave

**Valuation Gap**: US companies generally commanded higher P/E ratios,
reflecting market premiums for technological leadership

**Key Insight:** 2015 established a dual-track development pattern of
"US leading innovation, China leading application."

## Chapter 2: The Growth Race (2015-2018)

The mid-2010s witnessed explosive growth as mobile internet became
ubiquitous and platform business models scaled globally.

```{r growth-period, fig.height=7}

#2015-2018年增长期
growth_data <- cleaned_data %>%

filter(year(date) %in% 2015:2018)
#创建赛车图函数
create_race_chart <- function(data) {

p <- ggplot(data, aes(x = date, y = cumulative_return,

color = symbol, group = symbol,

text = paste("Company:", symbol,

"

Date:", date,

"

Return:", round(cumulative_return, 1), "%"))) +

geom_line(size = 1.2, alpha = 0.8) +

labs(

title = "Tech Giants (China vs US)",

subtitle = "Cumulative Return (Jan 2015 = 100)",

x = "Date",

y = "Return (%)",

color = "Company"

) +

scale_color_viridis_d() +

scale_x_date(date_breaks = "1 year", date_labels = "%Y") +

theme_minimal()

ggplotly(p, tooltip = "text")

}

create_race_chart(growth_data)
```

**Core Growth Phase Findings**:

### US Company Performance

-   **NVDA Extraordinary Growth**: Over 1500% cumulative returns,
    highlighting GPU's core value in the AI era
-   **Cloud Services Driven**: Microsoft and Amazon achieved secondary
    growth curves through cloud computing
-   **Platform Effects**: Google and Meta continuously expanded
    competitive advantages through network effects

### Chinese Company Characteristics

-   **Mobile Internet Dividend**: Tencent and Alibaba successfully
    migrated PC advantages to mobile platforms
-   **Localized Innovation**: Meituan and JD.com established barriers in
    O2O and e-commerce logistics
-   **Growth Quality**: Although growing faster, volatility was
    significantly higher than US counterparts

**Key Event Impact**: Cloud computing commercialization, smartphone
proliferation, and mobile payment explosion were the main drivers during
this phase.

## Chapter 3: Tests and Resilience (2018-2021)

This period challenged tech giants with trade tensions, pandemic
disruptions, and increasing regulatory scrutiny.

```{r challenge-period, fig.height=7}
#挑战期分析
challenge_data <- cleaned_data %>%

filter(year(date) %in% 2018:2021)

#创建热力图函数
create_returns_heatmap <- function(data) {

yearly_returns <- data %>%

mutate(year = year(date)) %>%

group_by(symbol, year) %>%

summarise(

yearly_return = (last(adjusted) - first(adjusted)) / first(adjusted),

.groups = 'drop'

)

p <- ggplot(yearly_returns, aes(x = year, y = symbol, fill = yearly_return)) +

geom_tile() +

labs(

title = "Heatmap of Return",

x = "Year",

y = "Company",

fill = "Return"

) +

scale_fill_gradient2(low = "red", mid = "white", high = "green",

midpoint = 0, labels = percent) +

scale_x_continuous(breaks = seq(2015, 2024, 1)) +

theme_minimal()

ggplotly(p)

}

create_returns_heatmap(challenge_data)
```

### Crisis Response Capability Analysis

**Trade Friction Period (2018-2019)**: - Chinese companies generally
faced greater pressure, with Baidu and Alibaba showing significant
drawdowns - US tech stocks demonstrated relative resilience, though
companies with China-dependent supply chains were affected

**Pandemic Impact and Recovery (2020-2021)**:

\- **Beneficiaries**: NVDA, Tesla and other "pandemic winners" performed
exceptionally

\- **Resilience Performance**: Tencent, Microsoft and other diversified
enterprises quickly adapted to remote work demands

\- **Recovery Divergence**: Chinese companies showed stronger policy
adaptability, while US companies further consolidated technological
advantages

**Heatmap Analysis**: 2020 emerged as the year with most significant
performance divergence, where digitalization levels determined corporate
risk resistance capabilities.

## Chapter 4: The New Landscape (2022-2024)

The most recent period has been defined by AI breakthroughs and
geopolitical realignments.

```{r current-era, fig.height=7}

#当前时代分析
current_data <- cleaned_data %>%

filter(year(date) >= 2022)

create_race_chart(current_data)
```

### AI Revolution and Structural Reshaping

**NVDA's Dominant Performance**: - Cumulative returns exceeded 30,000%,
reflecting the scarcity value of AI chips - Became a critical "enabler"
in US-China tech competition

**US-China Divergent Paths**:

\- **US Focus**: Foundation models and computing infrastructure (NVDA,
MSFT)

\- **China Focus**: Application scenario implementation (Tencent,
Baidu's AI applications)

**Regulatory Environment Impact**: - China's antitrust policies
temporarily suppressed valuations but fostered healthier competition -
US chip controls created new market barriers and opportunities

## Advanced Quantitative Analysis

This section provides deeper insights through sophisticated quantitative
techniques, moving beyond simple price trends to explore risk,
correlation, and market dynamics.

### Risk-Return Profile Assessment

Understanding the trade-off between risk and return is fundamental to
investment analysis. This chart plots annualized returns against
volatility, with bubble size representing total return over the decade.

```{r risk-return-analysis, fig.height=7}

create_risk_return_analysis(cleaned_data)
```

**Key Insights**:

\- **Efficient Frontier**: Companies positioned along the efficient
frontier offer the best risk-adjusted returns

\- **Volatility Clustering**: Chinese companies generally exhibit higher
volatility, reflecting different market structures

\- **Risk-Adjusted Performance**: Sharpe ratios help identify truly
superior performers after accounting for risk

### Correlation Structure

Markets don't move in isolation. This correlation matrix reveals how
closely these tech giants move together, providing insights into
diversification opportunities.

```{r correlation-analysis, fig.height=7}

create_correlation_analysis(cleaned_data)
```

**Patterns Observed**:

\- **Regional Clustering**: Stronger correlations within regions than
between them

\- **Sector Dynamics**: Companies in similar business segments show
higher correlation

\- **Diversification Benefits**: Low or negative correlations indicate
potential hedging opportunities

```{r regime-analysis, fig.height=7}

create_market_regime_analysis(cleaned_data)
```

**Regime Characteristics**:

\- **Growth Period (2015-2017)**: Mobile internet and cloud computing
drove returns

\- **Trade War (2018-2020)**: Supply chain disruptions created winners
and losers

\- **Pandemic & Recovery (2020-2021)**: Digital transformation
accelerated -

-**Inflation & AI Boom (2022-2025)**: AI infrastructure became the
dominant theme

### Momentum Effects

Momentum investing is a well-documented phenomenon. This analysis
examines whether past returns predict future performance.

```{r momentum-analysis, fig.height=7}

create_momentum_analysis(cleaned_data)
```

**Momentum Insights**:

\- **Persistence Patterns**: Evidence of momentum continuation in
certain periods

\- **Regional Differences**: Potential variation in momentum
effectiveness across markets

\- **Trading Implications**: Supports tactical momentum-based strategies

## Quantitative Conclusion

### Portfolio Construction Implications

1.  **Optimal Weighting**: Risk-return profiles inform strategic
    portfolio allocation
2.  **Hedging Strategies**: Correlation analysis suggests natural
    hedging pairs
3.  **Tactical Timing**: Market regime analysis supports dynamic
    allocation approaches

### Investment Strategy Applications

-   **Factor Investing**: Momentum and quality factors show predictive
    power
-   **Risk Parity**: Volatility targeting can improve risk-adjusted
    returns
-   **Regime Detection**: Early identification of market shifts creates
    alpha opportunities

### Limitations and Future Research

While comprehensive, this analysis has limitations including
survivorship bias, look-ahead bias, and the challenge of capturing all
relevant risk factors. Future research could incorporate additional
factors like valuation metrics, sentiment indicators, and macroeconomic
variables.

## Conclusion: Key Insights

### Core Findings Validation

1.  **Technology Cycle Dominance**: The AI revolution is replicating
    mobile internet era investment logic
2.  **Geopolitical Normalization**: Tech industry deeply intertwined
    with national security, global models facing restructuring
3.  **Niche Differentiation**: US strength in fundamental innovation,
    China strength in scale application

### Investment Implications

-   **Innovation Density** determines long-term value: NVDA's case
    proves technological barriers matter more than scale
-   **Policy Adaptability** becomes core competitiveness: US and Chinese
    companies must find growth paths in different regulatory
    environments
-   **Technology Convergence Trend**: Cloud-native and AI-native
    companies are reshaping traditional industry structures

### Methodological Value

This analysis validates the effectiveness of public market data-based
tech industry analysis frameworks, providing quantitative tools for
understanding complex technology-market-policy interactions.

## Methodology and Data

### Data Sources

-   **Primary Data**: Yahoo Finance via tidyquant package
-   **Period**: January 2015 - Current
-   **Companies**: Leading US and Chinese technology firms

### Analytical Approach

-   All returns calculated from adjusted closing prices
-   Cumulative returns normalized to 100 at starting period
-   Interactive visualizations created with Plotly

------------------------------------------------------------------------

*This analysis was completed as part of DATA5002 Data Visualization
course.*
